---
jupyter:
  jupytext:
    formats: ipynb,Rmd
    text_representation:
      extension: .Rmd
      format_name: rmarkdown
      format_version: '1.2'
      jupytext_version: 1.11.0
  kernelspec:
    display_name: R
    language: R
    name: ir
---

# FDA

---


## Function-on-scalar regression model


**Question of interest :** Is there any relationship between the weight loss and the activity counts?


We can use a scalar-on-function regression model to answer this question:  
$$Y_i = \mu_Y + \int{Act_i(t)\beta(t)} +\epsilon_i(t)$$

where $Act_i(t)$ is modeled using fPCA, $Act_i(t) = \mu(t) + \sum_{k=1}^\infty{\xi_{ik}\phi_k(t)}$


## Load input: packages, external functions, input data.


### Packages

```{r}
# Libraries
library(fda)
library(fdANOVA)
library(tidyverse)
library(dplyr)
library(refund)
library(refund.shiny)
#theme_set(theme_bw() + theme(legend.position = "bottom"))
library(ggplot2)
library(fields)
library(mgcv)
```

### Data

```{r}
# Read cvs file
# df = read_csv("../../../data/fda/mBaselineSteps60.csv");
# df = read_csv("../../../data/fda/mVisitBaseline_60_Acc.csv");
# dfAcc = read_csv("../../../data/fda/mVisit32w_60_Acc.csv");
# df = read_csv("../../../data/fda/mVisit6wPP_60_Acc.csv");
# df = read_csv("../../../data/fda/mVisit6mPP_60_Acc.csv");
# df = read_csv("../../../data/fda/mVisit12mPP_60_Acc.csv")
df = read_csv("../../../data/fda/mVisit32w-VisitBaseline_60_Acc.csv");
dfWeight = read_csv("../../../data/fda/weightChange21.csv");


names(dfAcc)[names(dfAcc)=="studyID"] <- "id"
dfAcc$id <- as.character(dfAcc$id)
head(dfAcc, 3)

names(dfWeight)[names(dfWeight)=="studyID"] <- "id"
dfWeight$id <- as.character(dfWeight$id)
head(dfWeight, 3)

```

Joint data frame according to dfWeight:

```{r}
df <- left_join(dfWeight, dfAcc, by = "id")
head(df, 3)
```

```{r}
filter(df, id == "151" )
```

## Functional principal components (FPC) regression


We fit functional linear regression by assuming $\beta(t) = \sum_{k=1}^\infty{\beta_k\phi_k(t)}$


### Define response and covariate:

```{r}
# Create temporal axis depending on the resolution of the data
time <- 1:(dim(df)[2]-3)
n <- dim(df)[1]

X <- as.matrix(df[,c(4:dim(df)[2])]) # functional covariate: physical activity
Y <- as.vector(df$weightRatio) # scalar response: weight change

# myDat <- data.frame(X = X)
# myDat$Y <- Y

dim(X);length(Y)
```

### Run fPCA on the functional covariate using fpca.sc in refund package.
`fpca.sc` decomposes functional observations using functional principal components analysis. A mixed model framework is used to estimate scores and obtain variance estimates.

```{r}
fpca_res <- fpca.sc(X, argvals = time, pve = 0.99) # pve is the proportion of variance explained: used to choose the number of principal components.

m <- length(time)
efn <- fpca_res$efunctions*sqrt(m) # estimated eigenfunctions
evals <- fpca_res$evalues/m # estimated eigen values 
scr <- fpca_res$scores/sqrt(m) # estimated FPC scores
npc <- fpca_res$npc # number of principal component needed to explain proportion pve of the variance in the observed curves 
```

```{r}
matplot(time, efn, type='l', lty=1, lwd=2,
        col=rainbow(fpca_res$npc),
        main = "Estimated eigenfunctions", ylab="",
        xlab = "time(hour)")
legend("topleft", col = rainbow(fpca_res$npc), 
       lwd=2, lty=1, legend = paste0("fPC", 1:fpca_res$npc),
       cex = 1.5, bty = "n")
```

```{r}
kPC <- 1
effect <- sqrt(evals[kPC])*efn[,kPC]
mu_hat <- fpca_res$mu # estimated mean function

par(mfrow = c(2,1))
plot(time, efn[,kPC], type='l', lwd=2, lty=1, 
     xlab = "time(hour)", ylab="", main = paste("fPC", kPC))

matplot(time, cbind(mu_hat-effect, mu_hat+effect), 
        col=c("red", "blue"), pch = c("-", "+"),
        main = paste0("fPC", kPC), 
        xlab = "time(hour)", ylab="")
lines(time, mu_hat, lwd=2)
```

### Multiple linear regression on the vector of scalar response weight change

```{r}
out = lm(Y ~ scr) ## Multiple linear regression
# summary(out)
```

```{r}
beta_hat = out$coefficients
beta_hat
```

Since fPCA selected the first 6 principal componenets based on the specified percentage of variance explained, we have 6 corresponding basis coefficients.


Let reconstruct the regression coefficient function:

```{r}
par(mfrow=c(1,1))
beta_fn_hat  = efn %*% as.matrix(beta_hat[-1], col=1)
plot(time, beta_fn_hat, type='l', lwd=2,
     xlab="time(hour)", ylab = "", main="estimated coefficient fn")
```

Let’s focus on looking at three random activity profiles:

```{r}
set.seed(123)
nCrv <- 3
selCrv <- sample(1:n, size = nCrv, replace = FALSE)
selCrv
```

```{r}
matplot(time, t(fpca_res$Yhat[selCrv,]), type='l', lwd=2, lty=1,
        xlab="time(hour)", ylab="Activity Count", main="Three randomly selected curves")
```

```{r}
par(mfrow=c(3,3))
for(i in 1:3){
    ind <- selCrv[i]
    demeaned <- fpca_res$Yhat[ind,] - as.vector(fpca_res$mu)
    
    matplot(time, t(fpca_res$Yhat[selCrv,] - t(matrix(rep(fpca_res$mu,3), nrow= m))), 
            type='l', lwd=2, lty=1, col = 'light grey',
        xlab="time(hour)", ylab="Activity Count (demeaned)", main="")
    lines(time, demeaned, type='l', lwd=2, col='red')
    

    plot(time, beta_fn_hat, type='l', lwd=2,
         xlab="time(hour)", ylab = "estimated coefficient fn", main="")
    plot(time, demeaned*beta_fn_hat,type='l', lwd=2, col='blue',
         xlab="time(hour)", ylab = "", ylim=c(-55, 70),
         main=round(mean(demeaned*beta_fn_hat), 2))
}
```

### Goodness-of-fit

```{r}
par(mfrow=c(1,1))
plot(Y, out$fitted, cex=0.5, ylab="Fitted", xlab="Observed")
abline(a = 0, b = 1)
```

```{r}
fpca_res <- fpca.sc(X, argvals = time, pve = 0.97)
Xhat <- fpca_res$Yhat
Yhat <- predict(fit, newdata = list(X = Xhat))

# goodness-of-fit
par(mfrow=c(1,1))
plot(Y, Yhat, cex=0.5, ylab="Fitted", xlab="Observed")
abline(a = 0, b = 1)
```

Calculate R-squared:

```{r}
Rsq = 1-sum((out$residuals)^2)/sum((Y- mean(Y))^2)
Rsq
```

## Functional linear model with mixed basis

```{r}
myDat <- data.frame(X, Y)
fit <- pfr(Y ~ lf(X, k = 10, bs = "ps", presmooth="bspline"), method = "REML", data = myDat)
coef <- coef(fit)


plot(coef$X.argvals, coef$value, ylab=expression(paste(beta(t))), 
     xlab="hour", xaxt = 'n', type='l', lwd=2,
     main="Estimated coefficient function")
axis(side = 1, at = seq(0,1,length.out = m), label=1:m)
```

```{r}
par(mfrow=c(3,1))
for(k in c(10, 15, 20)){
 fit <- pfr(Y ~ lf(X, k = k, bs = "cr"), method = "REML", data = myDat)
 coef <- coef(fit)


plot(coef$X.argvals, coef$value, ylab=expression(paste(beta(t))), 
     xlab="time", xaxt = 'n', type='l', lwd=2,
     main= paste0("Estimated coefficient function (", k, " basis fns)"))
axis(side = 1, at = seq(0,1,length.out = m), label=1:m)
 
}
```

```{r}
df[,c(2,3)] %>% group_by(arm)
```

```{r}

```
