---
jupyter:
  jupytext:
    formats: ipynb,Rmd
    text_representation:
      extension: .Rmd
      format_name: rmarkdown
      format_version: '1.2'
      jupytext_version: 1.11.0
  kernelspec:
    display_name: R
    language: R
    name: ir
---

# FDA

---


## Function-on-scalar regression model


**Question of interest :** What is the association between treatment assignment and activity pattern?


We can use a functiuon-on-scalar regression model to answer this question:  
$$Act_i(t) = \beta_0(t) + \beta_1(t)Treatment_i  +\epsilon_i(t)$$


## Load input: packages, external functions, input data.


### Packages

```{r}
# Libraries
library(fda)
library(fdANOVA)
library(tidyverse)
library(refund)
library(refund.shiny)
#theme_set(theme_bw() + theme(legend.position = "bottom"))
library(ggplot2)
library(fields)
library(mgcv)
```

### Data

```{r}
# Read cvs file
# df = read_csv("../../../data/fda/mBaselineSteps60.csv");
# df = read_csv("../../../data/fda/mVisitBaseline_60_Acc.csv");
# df = read_csv("../../../data/fda/mVisit32w_60_Acc.csv");
# df = read_csv("../../../data/fda/mVisit6wPP_60_Acc.csv");
df = read_csv("../../../data/fda/mVisit6mPP_60_Acc.csv");
# df = read_csv("../../../data/fda/mVisit12mPP_60_Acc.csv")
# df = read_csv("../../../data/fda/mVisit32w-VisitBaseline_60_Acc.csv");



names(df)[names(df)=="studyID"] <- "id"
df$id <- as.character(df$id)
head(df, 5)
```

```{r}
filter(df, id == "248" )
```

### Define response and predictor:

```{r}
# Create temporal axis depending on the resolution of the data
time <- 1:(dim(df)[2]-2)

Y <- as.matrix(df[,c(3:dim(df)[2])]) # response: physical activity
X <- as.vector(df$arm) # predictor: treatment assignment

myDat <- data.frame(X = X)
myDat$Y <- Y

dim(Y);length(X)
```

### Fit function-on-scalar regression model using the pffr function.
(The pffr function in the refund package can fit any functional linear model with functional response.)

```{r}
fit <- pffr(Y ~ X, data = myDat) # fit
yhat <- predict(fit, newdata = myDat) # get prediction according to the model
```

**Calculate $R^2$**   
$$R^2 = 1 - \frac{\sum_i{(y_i - \hat{y}_i)^2}}{\sum_i{(y_i - \overline{y}_i)^2}}$$

```{r}
Rsq_t <- 1-colSums((Y - yhat)^2) / colSums((Y - colMeans(Y))^2)
mean(Rsq_t)
```

```{r}
# set plot size
options(repr.plot.width=15, repr.plot.height=10)
```

```{r}
matplot(time, t(Y), type='l', lty = 1, 
        col='light grey', ylab="fitted", xlab="day")
matlines(time, t(yhat), type='l', lty = 1, lwd = 3)
```

```{r}
 rownames(Y) <- df$id

accel_df = as_refundObj(Y)
# value is the value of the activity count
# index is the percentage period over 24 hours
```

```{r}
ggplot(accel_df, aes(x = index, y = value, group = id)) + geom_line()
```

### Plot by group

```{r}
as_refundObj(Y) %>% 
    left_join(dplyr::select(df, id, arm)) %>%
    ggplot(aes(x = index, y = value, group = id, color = arm)) + geom_path(alpha = .1) +
    geom_smooth(aes(group = arm), se = FALSE)+
scale_color_gradientn(colours = rainbow(3))
```

### Correlation

```{r}
axtime <- c(0:23) #names(df)[c(3:dim(df)[2])]
```

```{r}
# set plot size
options(repr.plot.width=8, repr.plot.height=8)
# cor(Y)
# image(cor(Y))
image.plot(axtime, axtime, cor(Y), main='Correlation')
```

```{r}

```

```{r}
image.plot(c(axtime, axtime+24), c(axtime, axtime+24), cbind(rbind(cor(log(Y+1)), cor(log(Y+1))), rbind(cor(log(Y+1)), cor(log(Y+1)))), main='Correlation')
```

```{r}
Ysmth <- smAct(df, axtime)
```

```{r}
image.plot(axtime, axtime, cor(Ysmth), main='Correlation')
```

```{r}
# set plot size
options(repr.plot.width=8, repr.plot.height=8)
# cor(Y)
# image(cor(Y))
image.plot(axtime, axtime, cov(log(Y+1)), main='Covariance')
```

```{r}
image.plot(axtime, axtime, cov(Ysmth), main='Covariance')
```

## FANOVA

```{r}
# Order data frame with respect to the group label
df <- df[order(df$arm),]
head(df, 3)
```

```{r}
axtime
```

```{r}
# Select a time window
df <- df[,c(c(1,2),c(0:6)+3)]
```

```{r}
axtime <- c(0:6)
```

```{r}
# Define range for plot
intrng <- c(0,(dim(df)[2]-3))
```

```{r}
intrng
```

```{r}
# Extract activity matrix from data frame, add name and transpose
mat <- (as.matrix(df[,c(3:dim(df)[2])]))
rownames(mat) <- df$id
mat <- t(mat)
```

```{r}
# set plot size
options(repr.plot.width=15, repr.plot.height=10)
```

```{r}
plotFANOVA(x = mat, int = intrng)
```

```{r}
plotFANOVA(x = mat, group.label = as.character(df$arm), int= intrng)
```

```{r}
# set plot size
options(repr.plot.width=8, repr.plot.height=8)
```

```{r}
plotFANOVA(x = mat, group.label = as.character(df$arm), int= intrng, separately = TRUE)
```

```{r}
plotFANOVA(x = mat, group.label = as.character(df$arm), int= intrng, means = TRUE, smooth = T)
```

### Estimation of mean functions
Smooth each curve and take pointwise average.

```{r}
# Select data for visit = 1 and case = 1
dfAct1 <- subset(df, arm == 1)
Y1 <- as.matrix(dfAct1[,c(3:dim(df)[2])])
n1 <- length(unique(dfAct1$id))

# dfAct2 <- subset(df, arm == 2)
# n2 <- length(unique(dfAct2$id))

# dfAct3 <- subset(df, arm == 3)
# n3 <- length(unique(dfAct3$id))

```

```{r}
matplot(axtime, t(Y1), 
        type='l', lty=1, col="light grey",
        main = "Activity group 1",
        xlab="time", ylab="PA")

selIndiv <- sample(1:n1, size = 3, replace = FALSE)
selIndiv

matlines(axtime, t(Y1[selIndiv,]), 
         type='l', lty=1, lwd=2, col = rainbow(3))
```

```{r}
table(df[,2])
```

```{r}
dim(dfAct1)
```

```{r}
# head(dfAct1)
n = 9 # 13
minArm1 <- min(dfAct1[,n])
minArm1
which(dfAct1[,n] == minArm1)
dfAct1[37, 1]
```

```{r}
which(colnames(mat) == "591")
```

```{r}
matCut <- mat[,-37] 
```

```{r}
dfCut <- df[-37,] 
```

```{r jupyter={'source_hidden': True, 'outputs_hidden': True}}
# Create a matrix of zeros with the same dimension of Y1 
smoothActCurves1 <- array(0, dim(Y1))
n <- nrow(Y1)
for(j in 1:n1){
  # j = 1
    # smooth estimator using penalized cubic B-spline basis functions
   fit <- gam(Y1[j,] ~ s(axtime, k = 7, bs = 'cr'), # k = 10
             method = "REML") # gam() is used for smoothing
  # plot(tract, DTI.baseline$cca[j,])
  # lines(tract, fit$fitted)
  smoothActCurves1[j,] <- fit$fitted
}
```

```{r}
# Plot the smooth fit versus original
matplot(axtime, t(Y1[selIndiv,]), 
         type='l', lty=3, lwd=1, col = rainbow(3))
matlines(axtime, t(smoothActCurves1[selIndiv,]), 
         type='l', lty=1, lwd=1, col = rainbow(3))
```

```{r}
# Average smooth curves
meanHat1 <- colMeans(smoothActCurves1)
# Plot original data with the mean estimation
matplot(axtime, t(Y1), 
        type='l', lty=1, col="light grey",
        main = "Activity: group 1",
        xlab="tract", ylab="PA")
lines(axtime, meanHat1, col='blue', lwd=2)
```

```{r}
meanHat2 <-smActGrp(df, 2, axtime)
meanHat3 <-smActGrp(df, 3, axtime)
```

```{r}
diffMean <- tibble(
    time =  c(0:(dim(df)[2]-3)),
    diff12 = abs(meanHat2-meanHat1),
    diff13 = abs(meanHat3-meanHat1),
    diff23 = abs(meanHat3-meanHat2)
)
```

```{r}
ggplot(diffMean, aes(x=time)) + 
  geom_line(aes(y = diff12), color = "darkred", size=2) + 
  geom_line(aes(y = diff13), color="steelblue", size=2) +
  geom_line(aes(y = diff23), color= "seagreen4", size=2) +
  labs(y="Group Difference", x = "Time (hour)", color = "Legend") +
    scale_color_manual(values = colors)
```

### FANOVA Test

```{r}
set.seed(123)
```

```{r}

rownames(Ysmth) <- df$id
Ysmth <- t(Ysmth)
```

```{r}
fanova <- fanova.tests(x = Ysmth, group.label = df$arm)
```

```{r}
# fanova <- fanova.tests(x = matCut, group.label = dfCut$Arm)
fanova <- fanova.tests(x = mat, group.label = df$arm)
```

```{r}
fanova
```

```{r}
plotFANOVA(x = mat[c(1:6),], group.label = as.character(df$Arm), int= c(0, 6), means = TRUE, smooth = T)
```

##### fanovares <- fanova.tests(x = mat[c(8:20),], group.label = df$Arm)

```{r}
# Create temporal axis depending on the resolution of the data
time <- 1:(dim(df)[2]-2)

Y <- as.matrix(df[,c(3:dim(df)[2])]) # response: physical activity
X <- as.vector(df$Arm) # predictor: treatment assignment

myDat <- data.frame(X = X)
myDat$Y <- Y

dim(Y);length(X)
```

```{r}

```

```{r}
temp_df = df
temp_df$Arm = as.factor(temp_df$Arm)
```

```{r}
temp_df$accel = Y
```

```{r}
temp_df$mean_accel = apply(Y, 1, mean)
```

```{r}
mod1 = lm(mean_accel~Arm, data = temp_df)
summary(mod1)
```

```{r}
# kt number of basis functions
fosr_slr = bayes_fosr(accel ~ Arm, data = temp_df, 
                                            est.method = "GLS", Kt = 8, basis = "pbs")


## Warning in model.matrix.default(mt_fixed, mf_fixed, contrasts): non-list
## contrasts argument ignored
## Using OLS to estimate residual covariance 
## GLS

as_refundObj(fosr_slr$beta.hat) %>% 
    ggplot(aes(x = index, y = value)) + facet_grid(~id) + geom_path()
```

```{r jupyter={'source_hidden': True}}
> wheat = read.table('http://www.stat.berkeley.edu/classes/s133/data/wheat.txt',header=TRUE)
> plot(wheat$year,wheat$production)
> lines(smooth.spline(wheat$year,wheat$production))
```

## Functions


### Smooth mean function

```{r}

smActGrp <- function(dfAct, grp, timeAx){
    # Select data for visit = 1 and case = 1
    dfGrp <- subset(dfAct, arm == grp)
    Y <- as.matrix(dfGrp[,c(3:dim(dfAct)[2])])
    n <- length(unique(dfGrp$id))

    # Create a matrix of zeros with the same dimension of Y1 
    smoothActCurves <- array(0, dim(Y))
    
    for(j in 1:n){
      # j = 1
        # smooth estimator using penalized cubic B-spline basis functions
       fit <- gam(Y[j,] ~ s(timeAx, k = 7, bs = 'cr'), # k =10
                 method = "REML") # gam() is used for smoothing
      # plot(tract, DTI.baseline$cca[j,])
      # lines(tract, fit$fitted)
      smoothActCurves[j,] <- fit$fitted
    }

    # Average smooth curves
    meanHat <- colMeans(smoothActCurves)
    
    return(meanHat)
}
```

```{r}
test <-smActGrp(df, 1, axtime)

# Plot original data with the mean estimation
matplot(axtime, t(Y1), 
        type='l', lty=1, col="light grey",
        main = "Activity: group 1",
        xlab="tract", ylab="PA")
lines(axtime, test, col='blue', lwd=2)
```

### Smooth Activity

```{r}
smAct <- function(dfAct,timeAx){
    # Select data for visit = 1 and case = 1
    Y <- as.matrix(dfAct[,c(3:dim(dfAct)[2])])
    n <- length(unique(dfAct$id))

    # Create a matrix of zeros with the same dimension of Y1 
    smoothActCurves <- array(0, dim(Y))
    
    for(j in 1:n){
      # j = 1
        # smooth estimator using penalized cubic B-spline basis functions
       fit <- gam(Y[j,] ~ s(timeAx, k = 10, bs = 'cr'),
                 method = "REML") # gam() is used for smoothing
      # plot(tract, DTI.baseline$cca[j,])
      # lines(tract, fit$fitted)
      smoothActCurves[j,] <- fit$fitted
    }
   
    return(smoothActCurves)
}
```

```{r}
# mat <- (as.matrix(df[,c(3:dim(df)[2])]))
```

```{r}
dftmp <- data.frame( id= df[,1],
          Arm = df[,2],
          act = rowSums(df[,c(3:dim(df)[2])])
    )
```

```{r}
names(dftmp)
```

```{r}

rslt <- aov(act ~ as.factor(Arm), data = dftmp)

summary(rslt)
```

```{r}
mattmp <- (as.matrix(df[,c(3:dim(df)[2])]))
mattmp
```

```{r}
vec1 <- c(1,1,2,3)
```

```{r}
vec1
```

```{r}
unique(vec1)
```

```{r}
length(vec1)
length(unique(vec1))
```

```{r}

```
