---
jupyter:
  jupytext:
    formats: ipynb,Rmd
    text_representation:
      extension: .Rmd
      format_name: rmarkdown
      format_version: '1.2'
      jupytext_version: 1.5.2
  kernelspec:
    display_name: R
    language: R
    name: ir
---

# FDA

---


## Functional linear models


### Function-on-scalar regression model


**Question of interest :** What is the association between treatment assignment and activity pattern?


We can use a functiuon-on-scalar regression model to answer this question:  
$$Act_i(t) = \beta_0(t) + \beta_1(t)Treatment_i  +\epsilon_i(t)$$

$$  \sum_{G = A, B, C}{\frac{n_G(\overline{G}-\overline{X})^2}{N}}$$


**Load libraries and read the data:**

```{r}
# Libraries
library(fda)
library(tidyverse)

library(refund)
library(refund.shiny)

# theme_set(theme_bw() + theme(legend.position = "bottom"))

library(ggplot2)
```

```{r}
getwd()
```

```{r}
# Read cvs file
df = read_csv("../../../data/fda/mBaseline.csv");
names(df)[names(df)=="StudyID"] <- "id"
df$id <- as.character(df$id)
head(df, 5)
```

**Define response and predictor:**

```{r}
time <- 1:(dim(df)[2]-2)
time
```

```{r}
# Create temporal axis depending on the resolution of the data
time <- 1:(dim(df)[2]-2)

Y <- as.matrix(df[,c(3:dim(df)[2])]) # response: physical activity
X <- as.vector(df$Arm) # predictor: treatment assignment

myDat <- data.frame(X = X)
myDat$Y <- Y

dim(Y);length(X)
```

**Fit function-on-scalar regression model using the pffr function.** (The pffr function in the refund package can fit any functional linear model with functional response.)

```{r}
# library(refund)
```

```{r}
fit <- pffr(Y ~ X, data = myDat) # fit
yhat <- predict(fit, newdata = myDat) # get prediction according to the model
```

**Calculate $R^2$**   
$$R^2 = 1 - \frac{\sum_i{(y_i - \hat{y}_i)^2}}{\sum_i{(y_i - \overline{y}_i)^2}}$$

```{r}
Rsq_t <- 1-colSums((Y - yhat)^2) / colSums((Y - colMeans(Y))^2)
mean(Rsq_t)
```

```{r}
matplot(time, t(Y), type='l', lty = 1, 
        col='light grey', ylab="fitted", xlab="day")
matlines(time, t(yhat), type='l', lty = 1)
```

```{r}
 rownames(Y) <- df$id
```

```{r}
accel_df = as_refundObj(Y)
# value is the value of the activity count
# index is the percentage period over 24 hours
```

```{r}
ggplot(accel_df, aes(x = index, y = value, group = id)) + geom_line()
```

```{r}
# set plot size
options(repr.plot.width=15, repr.plot.height=10)
```

```{r}
as_refundObj(Y) %>% 
    left_join(dplyr::select(df, id, Arm)) %>%
    ggplot(aes(x = index, y = value, group = id, color = Arm)) + geom_path(alpha = .1) +
    geom_smooth(aes(group = Arm), se = FALSE)+
scale_color_gradientn(colours = rainbow(3))
```

```{r}
# set plot size
options(repr.plot.width=8, repr.plot.height=8)
cor(Y)
image(cor(Y))
```

## FANOVA

```{r}
library(fdANOVA)
```

```{r}
# Order data frame with respect to the group label
df <- df[order(df$Arm),]
head(df, 3)
```

```{r}
mat <- (as.matrix(df[,c(3:dim(df)[2])]))
rownames(mat) <- df$id
mat <- t(mat)
```

```{r jupyter={'outputs_hidden': True}}
plotFANOVA(x = mat, int = intrng)
```

```{r}
plotFANOVA(x = mat, group.label = as.character(df$Arm), int= intrng)
```

```{r}
plotFANOVA(x = mat, group.label = as.character(df$Arm), int= intrng, separately = TRUE)
```

```{r}
plotFANOVA(x = mat, group.label = as.character(df$Arm), int= intrng, means = TRUE, smooth = T)
```

```{r}
set.seed(123)
```

```{r}
fanova <- fanova.tests(x = mat, group.label = df$Arm)
```

```{r}
fanova
```

```{r}
plotFANOVA(x = mat[c(8:20),], group.label = as.character(df$Arm), int= c(0, 23), means = TRUE, smooth = T)
```

```{r}
fanovares <- fanova.tests(x = mat[c(8:20),], group.label = df$Arm)
```

```{r}
# Create temporal axis depending on the resolution of the data
time <- 1:(dim(df)[2]-2)

Y <- as.matrix(df[,c(3:dim(df)[2])]) # response: physical activity
X <- as.vector(df$Arm) # predictor: treatment assignment

myDat <- data.frame(X = X)
myDat$Y <- Y

dim(Y);length(X)
```

```{r}
head(df)
```

```{r jupyter={'source_hidden': True}}
gait.data.frame <- as.data.frame(gait)
```

```{r jupyter={'source_hidden': True, 'outputs_hidden': True}}
head(gait.data.frame)
```

```{r}
x.gait <- vector("list", 2)
```

```{r jupyter={'source_hidden': True, 'outputs_hidden': True}}
x.gait
```

```{r jupyter={'source_hidden': True, 'outputs_hidden': True}}
x.gait[[1]] <- as.matrix(gait.data.frame[, 1:39])
x.gait
```

```{r jupyter={'source_hidden': True}}
group.label.gait <- rep(1:3, each = 13)
```

```{r jupyter={'source_hidden': True, 'outputs_hidden': True}}
group.label.gait
```

```{r}

```

```{r}

```

```{r}

```

```{r}
temp_df = df
temp_df$Arm = as.factor(temp_df$Arm)
```

```{r}
temp_df$accel = Y
```

```{r}
temp_df$mean_accel = apply(Y, 1, mean)
```

```{r}
mod1 = lm(mean_accel~Arm, data = temp_df)
summary(mod1)
```

```{r}
# kt number of basis functions
fosr_slr = bayes_fosr(accel ~ Arm, data = temp_df, 
                                            est.method = "GLS", Kt = 8, basis = "pbs")


## Warning in model.matrix.default(mt_fixed, mf_fixed, contrasts): non-list
## contrasts argument ignored
## Using OLS to estimate residual covariance 
## GLS

as_refundObj(fosr_slr$beta.hat) %>% 
    ggplot(aes(x = index, y = value)) + facet_grid(~id) + geom_path()
```

```{r}
tt = tibble(
    A = c(1,1,1,2,2,2,3,3,3)
    )
```

```{r}
dd <- tibble(
    A = c(1,2,3),
    B = c("a","b","c")
    )
```

```{r}
dd
```

```{r}
left_join(tt,dd)
```

```{r}
> wheat = read.table('http://www.stat.berkeley.edu/classes/s133/data/wheat.txt',header=TRUE)
> plot(wheat$year,wheat$production)
> lines(smooth.spline(wheat$year,wheat$production))
```
